---
output:
  html_document: default
  pdf_document: default
---

STAT 1223 Project Statistical Analysis

Packages

```{r}
library(tidyverse)
library(zoo)
library(lmtest)
```

Load Dataset

```{r}
Treasury <- read.csv("Merged_Data.csv", sep=",", header = TRUE)
```

Clean Data

```{r}
  # Convert date
Treasury$date <- as.Date(Treasury$Date)

str(Treasury)

  # Create original dataset without unnecessary Date columns, ready to use to model
Treasury_original <- Treasury %>% select(-Date, -date)

  # COVID dummy
Treasury_original$covid <- ifelse(as.Date(Treasury$date) >= as.Date("2020-03-01") &
  as.Date(Treasury$date) <= as.Date("2021-06-01"), 1, 0)

  # SP500 change (first-difference)
Treasury_original$SP500_diff <- c(NA, diff(Treasury$SP500))

 # Lagged variable (inflation)
Treasury_original$inflation_lag2 <- dplyr::lag(Treasury_original$Inflation_Rate, 2)
```

Drop NA

```{r}
treasury_na <- Treasury_original %>% drop_na(inflation_lag2, SP500_diff)
```

5-number summary + mean + histogram

```{r}
summary(treasury_na$US_10Y)

mean(treasury_na$US_10Y, na.rm = TRUE)
sd(treasury_na$US_10Y, na.rm = TRUE)

png("us10y_histogram2.png", width = 800, height = 600)
hist(Treasury$US_10Y,
     main = "Histogram of US10Y",
     xlab = "US10Y",
     col = "skyblue")
dev.off()
```

```{r}
summary(treasury_na %>% select(FEDFUNDS_RATE, UNRATE, GOLD, USD, CN10Y, inflation_lag2, SP500_diff))

sapply(treasury_na %>% select(FEDFUNDS_RATE, UNRATE, GOLD, USD, CN10Y, inflation_lag2, SP500_diff),
       sd, na.rm = TRUE)
```

PLOTS: Check for linearity

```{r}
library(ggplot2)
```

```{r}
Treasury_long1 <- treasury_na %>%
  pivot_longer(cols = -c(US_10Y), names_to = "Predictor", values_to = "Value")
```

```{r}
predictor_names1 <- unique(Treasury_long1$Predictor)
```

```{r}
for (var in predictor_names1) {
  Treasury_plot_data <- Treasury_long1 %>%
    filter(Predictor == var)
   scale_factor <- max(Treasury_plot_data$Value, na.rm = TRUE) / max(Treasury$US_10Y, na.rm = TRUE)
  
  p <- ggplot(Treasury_plot_data, aes(x = date, y = Value)) +
    geom_line(aes(y = Value), color = "red", size = 1) +
    
    # Right y-axis: US_10Y time series (rescaled)
    geom_line(aes(y = US_10Y * scale_factor), color = "blue", size = 1) +
    labs(title = paste("Time-Based Relationship: US_10Y vs", var),
         x = "Date",
         y = var) +
      
    scale_y_continuous(
      sec.axis = sec_axis(~ . / scale_factor, name = "US_10Y")  # Convert back to original US_10Y scale
    ) +
    
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5, size = 16),
          axis.title = element_text(size = 14),
          axis.text = element_text(size = 12))

    print(p)  # Print each plot separately
}
```
```{r}
for (var in predictor_names1) {
  
  p <- ggplot(Treasury_long1 %>% filter(Predictor == var), aes(x = Value, y = US_10Y)) +
    geom_point(alpha = 0.5) +  # Scatter points
    geom_smooth(method = "lm", color = "red", se = TRUE) +  # Linear regression line
    geom_smooth(method = "loess", color = "blue", se = FALSE) +  # LOESS curve (for comparison)
    labs(title = paste("Linearity Check: US_10Y vs", var),
         x = var,
         y = "US_10Y") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5, size = 16),
          axis.title = element_text(size = 14),
          axis.text = element_text(size = 12))

  print(p)
 
ggsave(filename = paste0("scatterplot_US10Y_vs_", var, ".png"),
       plot = p,
       width = 6,
       height = 4,
       dpi = 300)
}
```

Box Plot

```{r}
p1 <- ggplot(treasury_na, aes(x = factor(covid),
                               y = US_10Y,
                               fill = factor(covid))) +
  geom_boxplot(outlier.color = "darkred", outlier.shape = 16, alpha = 0.7) +
  scale_x_discrete(labels = c("0" = "Non-COVID", "1" = "COVID")) +
  labs(title = "Figure 8. US10Y Yields During and Outside COVID Periods",
       x = "Period",
       y = "US 10-Year Treasury Yield (%)") +
  theme_minimal() +
  theme(legend.position = "none")

print(p1)

ggsave("us10y_covid_boxplot.png", plot = p1, width = 6, height = 4, dpi = 300)

```

Collinearity check

```{r}
# Correlation Matrix
cor(Treasury_original %>% select(US_10Y, FEDFUNDS_RATE,GOLD, Inflation_Rate, SP500, UNRATE, USD))
```

```{r}
cor_matrix1 <- cor(treasury_na %>% select(US_10Y, FEDFUNDS_RATE,GOLD, inflation_lag2, SP500_diff, UNRATE, USD))
write.csv(cor_matrix1, "correlation_matrix.csv")

cor_matrix2 <-cor(treasury_na %>% select(US_10Y, FEDFUNDS_RATE,CN10Y, GOLD, inflation_lag2, SP500_diff, UNRATE, USD))
write.csv(cor_matrix2, "correlation_matrix2.csv")
```

Collinearity check after dropping problematic variables

```{r}
cor_data <- treasury_na %>% select(US_10Y, FEDFUNDS_RATE, GOLD, SP500_diff, USD, UNRATE, inflation_lag2)

cor(cor_data)
```

Run initial regression

```{r}
model <- lm(US_10Y ~ FEDFUNDS_RATE + USD + GOLD + UNRATE + SP500_diff + inflation_lag2 + covid, data = treasury_na)
summary(model)
```

VIF Check

```{r}
require(regclass)
VIF(model)
```

Durbin-Watson Test for autocorrelation

```{r}
dwtest(model)
```
Fix Autocorrelation

```{r}
library(sandwich)
```

```{r}
coeftest(model, vcov = NeweyWest(model))
```

Consider adding interaction term on Unemployment rate

```{r}
model1 <- lm(US_10Y ~ FEDFUNDS_RATE + USD + GOLD + UNRATE * covid + SP500_diff + inflation_lag2, data = treasury_na)
summary(model1)
```

```{r}
coeftest(model1, vcov = NeweyWest(model1))
```

```{r}
dwtest(model1)
```
```{r}
plot(model1, which = 1)

png("residuals_vs_fitted.png", width = 800, height = 600)
plot(model1, which = 1)
dev.off()
```
```{r}
library(apaTables)

apa.reg.table(model1, filename = "regression_final_table.doc")

```

